### 2016-07-08 capability-machine.tex
Added heap satisfaction monotonicity lemma
added mono lemmas for read/write condition, added read cond to entry cond, changed updatePcPerm to be total

\subsection*{Semantics}
Assume a $\decode$ function that decodes integer to instructions:
\decode &:\ints \rightarrow \Instrs
Assume an $\encodePerm$ function that encodes a permission as an integer:
\encodePerm &: \Perms \rightarrow \ints
  \Phi & \rightarrow \sem{\decode(\memreg(\pcreg))}(\Phi) & &                                   
                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
                                                                \quad\text{and $\perm \in \{ \exec,\rwx \}$ }
\Phi & \rightarrow \failed                                 & & \text{otherwise}
It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result. As we will see in the semantics, it is assumed that the initial process starts with a readwrite capability for this location.
  \executeAllowed{\perm} &=
                             \true & \text{if } \perm \in \{ \rwx, \exec, \entry \} \\
  \readAllowed{\perm} &=
                             \true & \text{if } \perm \in \{ \rwx, \exec, \readwrite, \readonly \} \\
  \writeAllowed{\perm} &=
                             \true & \text{if } \perm \in \{ \rwx, \readwrite\} \\
  \updatePcPerm{(\perm,\start,\addrend,\addr)} &=
                                       (\perm,\start,\addrend,\addr) & \text{if $\perm\in\{ \rwx, \exec \}$} \\
                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}

@@

\begin{align*}
  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
\end{align*}

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] }  \} 
\union \\
                   & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\exec]} \mid \\
                   & \quad\readCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \}
\union \\
% For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
                   & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\rwx]} \mid \\
                   & \quad \writeCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\rwx} \}
\end{align*}

@@

``Continuation'' relation:
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: Check if this is still monotone
  \stdkr(W) \defeq & \{ \npair{c} \mid c \in \stdvr(W) \land \\
                   & \quad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat{n}{W'}, \heap_f \ldotp \forall \reg \in \stdrr(W') \\
                   & \qquad \npair{(reg\update{\pcreg}{\updatePcPerm{\var{c}}}, \heap \uplus \heap_f} \in \observations(W') \}
\end{align*}
We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.

@@

>>>>

\subsection*{Semantics}
Assume a $\decode$ function that decodes integer to instructions:
\begin{align*}
\decode &:\ints \rightarrow \Instrs
\end{align*}
Assume an $\encodePerm$ function that encodes a permission as an integer:
\begin{align*}
\encodePerm &: \Perms \rightarrow \ints
\end{align*}
\begin{align*}
  \Phi & \rightarrow \sem{\decode(\memreg(\pcreg))}(\Phi) & &                                   
                                                              \arraycolsep=0pt
                                                              \begin{array}{l}
                                                                \text{if $\memreg(\pcreg) = \stdcap$}\\
                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
                                                                \quad\text{and $\perm \in \{ \exec,\rwx \}$ }
                                                              \end{array}\\
\Phi & \rightarrow \failed                                 & & \text{otherwise}
\end{align*}
It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result. As we will see in the semantics, it is assumed that the initial process starts with a readwrite capability for this location.
\begin{align*}
  \executeAllowed{\perm} &=
                           \begin{cases}
                             \true & \text{if } \perm \in \{ \rwx, \exec, \entry \} \\
                             \false & \text{otherwise}
                           \end{cases} \\
  \readAllowed{\perm} &=
                           \begin{cases}
                             \true & \text{if } \perm \in \{ \rwx, \exec, \readwrite, \readonly \} \\
                             \false & \text{otherwise}
                           \end{cases} \\
  \writeAllowed{\perm} &=
                           \begin{cases}
                             \true & \text{if } \perm \in \{ \rwx, \readwrite\} \\
                             \false & \text{otherwise}
                           \end{cases} \\
  \updatePcPerm{(\perm,\start,\addrend,\addr)} &=
                                     \begin{cases}
                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}
                                       (\perm,\start,\addrend,\addr) & \text{otherwise} \\

@@

\begin{align*}
  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \land \\
                                          & \; \quad \readCond{n',W',\addrend,\addr}
\end{align*}

\begin{lemma}[Write condition implies read condition]
\[
\forall n, W, \start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
\]
\end{lemma}
\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] }  \} 
\union \\
                   & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\exec]} \mid \\
                   & \quad\readCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \}
\union \\
% For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
                   & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\rwx]} \mid \\
                   & \quad \writeCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\rwx} \}
\end{align*}

@@

``Continuation'' relation:
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: Check if this is still monotone
  \stdkr(W) \defeq & \{ \npair{c} \mid c \in \stdvr(W) \land \\
                   & \quad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat{n}{W'}, \heap_f \ldotp \forall \reg \in \stdrr(W') \\
                   & \qquad \npair{(reg\update{\pcreg}{\updatePcPerm{\var{c}}}, \heap \uplus \heap_f)} \in \observations(W') \}
\end{align*}
We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.

@@

\begin{lemma}[Heap satisfaction downwards closure]
\[
  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
\]
\end{lemma}

\begin{lemma}[Read condition downwards closure]
\[
\forall n,n',W,\start, \addrend \ldotp \readCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \readCond{n',W,\start,\addrend}
\]
\end{lemma}
\begin{lemma}[Write condition downwards closure]
\[
\forall n,n',W,\start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \writeCond{n',W,\start,\addrend}
\]
\end{lemma}
